Domestication and the impacts of linked selection on the maize genome

### 2015-08-01 manuscript.tex
first forays

\documentclass{pnastwo}
\usepackage{cite}
\usepackage{color}
\usepackage{pgf}
%\usepackage{hyperref}
\usepackage[normalem]{ulem}
\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }
\newcommand{\citex}{\textcolor{blue}{\bf CITE}}
\newcommand{\X}{\textcolor{blue}{\bf X}}

@@

\title{Demographic history and patterns of selection in maize during and since domestication}
\author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
    California, Davis, CA, USA}\affil{2}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{3}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} Li Wang\affil{4}{Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun
  Durvasula\affil{1}{}, Matthew Hufford\affil{4}{}, \and Jeffrey
  Ross-Ibrarra\affil{1}{}\affil{5}{Genome Center and Center for population biology, University of
    California, Davis, CA, USA} }

\significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor a functional variant, have been carefully evaluated in only a few species. In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes. We show how demography interacts with selection to impact genome-wide patterns of diversity, including the important observation that rapid population expansion can increase the efficiency of selection as much as a sudden population bottleneck can weaken it. Along the way, we develop the first estimate the demographic parameters of the maize domestication from whole genome sequence data.}

@@

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby genetic diversity surrounding
that mutation is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous (specifically missense) and synonymous substitutions in maize and
teosinte since divergence from \emph{Tripsacum}, roughly 1-1.2 million years before
present \cite{ross2009}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified (Figure
\ref{fig:hardSweeps}). This observation suggests that hard sweeps are not a primary form of
selection for either maize or teosinte.

To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}. In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure S1), confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize. Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize. When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure S2).

\begin{figure*}[!htb]
%\vspace*{.05in}
\centering
\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_June}
\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_June}
\caption{Pairwise diversity surrounding synonymous and non-synonymous
  substitutions in maize and teosinte. {\bf A:} Maize diversity; {\bf B:} Teosinte
diversity. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
\end{figure*}

@@

\subsection{Pairwise diversity is heavily influenced by purifying selection}
% \jri{we will need to define background selection and be careful about ``purifying selection'' vs ``background selection'', as technically background selection does not include very weak deleterious alleles.}
Purifying selection refers to the situation where deleterious mutations arising in a population are continuously selected against.
When this form of selection is operating, it can serve to reduce genetic variability at linked neutral sites, a phenomenon often called background selection \cite{charlesworth1993}. Purifying and background selection lead to lower diversity within genes and other functional sites relative to neutral regions.
We investigated purifying selection in maize and teosinte by evaluating the average magnitude of reduced diversity within genes and recovery away from genes in both taxa.
To enable comparisons between maize and teosinte, we standardized by neutral levels within each taxa, where neutral levels were defined as mean diversity at positions at least 0.01 cM distal from genes. For both maize and teosinte, pairwise diversity was reduced within genes and a recovery of diversity away from genes was observed. Interestingly, a stronger reduction of diversity and slower recovery was observed for teosinte than for maize, implying that purifying selection has left a more pronounced signature on pairwise diversity in the teosinte genome (Figure \ref{fig:purify}).

@@

We used diffusion approximation as implemented in
dadi \cite{gutenkunst2009} to find the domestication parameters that best explain
the joint site frequency spectrum of maize and teosinte.  The model we optimized began with an
equilibrium ancestral population of size $N_a$
splitting into separate maize and teosinte populations $T_B$ generations in the past. Moving forward in
time from the split, teosinte maintained the ancestral population size, while
maize experienced an immediate effective population size change to
$N_b$ individuals, followed by exponential growth to size
$N_m$. Additionally, since the split $M_{tm}$ individuals migrated
from teosinte into maize and $M_{mt}$ individuals migrated from maize
to teosinte.

@@

\begin{figure}[!htb]
%\vspace*{.05in}
\centering
\includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
\caption{Parameters of domestication as estimated by dadi. Notably,
  the maize effective population size ($N_e$) during the domestication
  bottleneck appears to have consisted of approximately 5.26\%
  the ancestral $N_e$, before recovering two at least 2.98
  times as large as the ancestral $N_e$. \label{fig:bottleneck} }
\end{figure}

@@

Ultimately, we have shown that purifying selection in maize has operated very differently than purifying selection in teosinte. This observation, along with the potential for soft sweeps, appear to explain the phenotypic divergence between maize and teosinte much more completely than dooes positive selection generating hard-sweeps at protien-coding mutations. Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the deomography of crop domestication can impact crop diversity. The bottleneck-effects from a sudden collapse in population size have been well studied and are known to impact crops for thousands of generations. Complementing this knowledge, our results demonstrate that the rapid expansion experienced by many crops after domestication can also have a profound influence on patterns of diversity, and the effects of this expansion should be accounted for as important contributors to long-term evolution.

@@

\subsection{Plant materials}
Accessions studied were selected from the Maize HapMap2
panel \cite{chia2012, lemmon2014}. Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results (Figure S5). Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion. 
Thirteen teosinte inbred lines, all members of  subspecies Z. \emph{mays}
ssp. \emph{parviglumis}, were utilized. Included lines are depicted in (Table S1).
Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.

\subsection{Interpolating genetic position}
For most of our analyses, physical position along a chromosome was a less relevant measure of map location than genetic position.
Physical positions were converted to genetic positions by interpolating from the NAM genetic map \cite{glaubitz2014}, which provides a 1 cM resolution for physical to genetic conversion.
Within R \cite{R2014}, physical positions with corresponding genetic positions in the NAM map were used as anchors.
Physical positions in our dataset wihout corresponding genetic position were assigned genetic positions by scaling the anchored genetic positions according to the physical distance between the unlabeled position and the flanking anchors.

\subsection{Estimating the site frequency spectrum}
To estimate individual  and joint site frequency spectra (SFS) for maize and teosinte from inbred lines, each inbred individual was treated as representing a single haplotype from its population.
SFS were
separately computed for genic and intergenic regions, as well as for
the whole genome together. First, genic and intergenic regions were isolated using
the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014}. Genic
regions were defined as DNA between the start and stop position of a
gene, while intergenic regions were required to be at least 5kb up- or
down-stream from a gene start/stop. With regions defined, SFS were
estimated with ANGSD \cite{korneliussen2014}. Individual population SFS were
estimated using all positions observed in at least 80\% of the
individuals in the population, and joint SFS were estimated using all
positions observed in at least 80\% of individuals in both
populations. Individuals were
assumed to be fully inbred (-doSaf 2). Quality filters were
employed such that reads with quality score below 30 and bases with
quality score below 20 were discarded (-minMapq 30 and -minQ 20), as were reads that didn't map
uniquely (-uniqueOnly 0). Quality scores around indels were adjusted
as in Samtools (-baq 0). Genotype likelihoods were estimated using the samtools
method (-GL 1). Major and minor alleles were inferred from the data
(-doMaf 1). Because ANGSD cannot calculate a folded joint SFS, the
maize reference genome was used for polarization and then unfolded spectra
were folded using dadi \cite{gutenkunst2009}.

\subsection{Demographic inference}
We used dadi \cite{gutenkunst2009} to estimate the parameters of maize
domestication based on diffusion approximation of the 2-dimensional
SFS for maize and teosinte. Our first step was to specify a model, as
shown in (Figure \ref{fig:bottleneck}). In words, our model began with an ancestral teosinte
population with effective population size $N_a$ individuals. The
ancestral population split into
two distinct populations, one of maize and the other of teosinte,
$T_b$ generations in the past. The maize population immediately shrank
to size $N_b$, and then grew exponentially to size $N_{maize}$
today. From the split until today, the teosinte effective population
size remained constant. Each year, $M_{mt}$ individuals migrated from
maize to teosinte, and $M_{tm}$ individuals migrated from teosinte to
maize.The free parameters of our model were $N_a, N_b, N_{maize}, T_b,
M_{mt}, \text{and } M_{tm} $. We implimented 1,000 dadi
optimizations of this model using the ``dadi.Inference.optimize\_log\_fmin''
optimizer. In dadi optimizations,
initial values for $N_b, N_{maize}, T_b,
M_{mt}, \text{and } M_{tm}$, respectively, were randomly perturbed up
to a factor of two from 0.02, 3, 0.04, 1e-5, and 1e-5. Lower
bounds were set at 1e-7, 1e-7,  0, 1e-10, and
1e-10, respecively, while upper bounds were respectively set at 2,
200, 1, .001, and .001. $N_a$ did not have an initial value, upper, or
lower bounds because it is estimated from equilibrium ancestral theta
($\theta = 4N_a\mu$), which dadi determines automatically regardless
of model. Population sizes were converted from dadi's
units using a mutation rate estimate of 3e-8
\cite{clark2005}. Among the 1,000 dadi runs, parameters
from the run with the lowest log-likelihood were taken as estimates of
the parameters of the domestication process.

Recent population expansion, as is observed for maize, most influences
the abundance of rare alleles. However, our demographic parameter
estimation using dadi suffered from a relatively small sample size of
23 maize individuals. Therefore, our approach likely underestimated
current maize effective population size. To address this, we utilized
a complimentary dataset of 4,021 maize landrace individuals collected
from across the Americas \textcolor{red}{(SeeDs reference here)} to
generate an independent estimate of
$N_{maize}$. These individuals were genotyped using GBS
\cite{elshire2011}, a protocol that suffers from a high rate of
missing data \cite{beissinger2013}. To minimize biases imposed from
missing data, we used R to isolate only those SNPs that
were observed in at least 1,500 individuals and subsequently projected the
SFS down to a sample of 500 individuals. Then, we utilized the singleton-based estimate of
theta ($\theta = 4N_e\mu = \frac{S}{L}$) to estimate
modern maize $N_e$ \cite{fu1993}, where $S$ is the total number of
singletons and $L$ is the total length of DNA sequenced. Based on the
GBS protocol implemented \cite{elshire2011}, $L$ is equal to the total
number of SNPs in the dataset.

>>>>

\documentclass{pnastwo}
\usepackage{cite}
\usepackage{color}
\usepackage{pgf}
%\usepackage{hyperref}
\usepackage[normalem]{ulem}
\newcommand{\jri}[1]{\textcolor{red}{\emph{#1}} }
\newcommand{\citex}{\textcolor{red}{\bf CITE}}
\newcommand{\X}{\textcolor{red}{\bf X}}

@@

\title{Domestication and the impacts of linked selection on the maize genome}
\author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
    California, Davis, CA, USA}\affil{2}{US Department of Agriculture, Agricultural Research Service, Columbia, MO, USA}\affil{3}{Division of Plant Sciences, University of Missouri, Columbia, MO, USA} Li Wang\affil{4}{Iowa State University, Ames, IA, USA}, Kate Crosby\affil{1}{}, Arun
  Durvasula\affil{1}{}, Matthew Hufford\affil{4}{}, \and Jeffrey
  Ross-Ibarra\affil{1}{}\affil{5}{Genome Center and Center for population biology, University of
    California, Davis, CA, USA} }

\significancetext{Patterns of linked selection, or the impact of selection on sites that neighbor a functional variant, have been carefully evaluated in only a few species. \jri{not really. we need to cite the Corbett-Detig plos-bio paper here -- he looked at a lot of species including maize!} In this work, we demonstrate that selection against deleterious mutations leaves a pronounced signature on the maize genome, reducing diversity in and immediately around genes. We show how demography interacts with selection to impact genome-wide patterns of diversity, including the important observation that rapid population expansion can increase the efficiency of selection as much as a sudden population bottleneck can weaken it. Along the way, we develop the first estimate the demographic parameters of the maize domestication from whole genome sequence data. }

@@

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby surrounding genetic diversity
is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous (specifically missense) and synonymous substitutions in maize and
teosinte since divergence from \emph{Tripsacum}, roughly 1-1.2 million years before
present \cite{ross2009}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified (Figure
\ref{fig:hardSweeps}). This observation suggests that hard sweeps are not a primary form of
selection for either maize or teosinte.

To ensure that this analysis is resilient to the intricacies of our sample, we investigated diversity surrounding the maize gene \emph{tga1}, one of the few known instances of a hard sweep corresponding to a causal missense substitution in maize \cite{wang2015}. In the case of this known sweep, we observed a reduction in diversity surrounding the causative substitution to levels much lower than seen for synonymous substitutions (Figure S1) \jri{please add refs to link to supp. figs if possible}, confirming that this method should be informative if hard sweeps were the prevailing mode of selection in maize. Still, to ensure that the approach is robust to potential shortcomings such as limited power in the face of background selection \cite{enard2014}, we sought to determine if patterns of diversity surrounding synonymous and non-synonymous substitutions differ when we restrict our analysis to the most- or least-conserved genes in maize. When we re-analyzed subsets of genes with the 10\% highest and 10\% lowest genomic evolutionary rate profile (GERP) score, putatively corresponding to genes undergoing the strongest and weakest purifying selection \cite{davydov2010, cooper2005, rodgers2015}, we again saw no significant difference in diversity surrounding synonymous and non-synonymous sites (Figure S2).

\begin{figure*}[!htb]
%\vspace*{.05in}
\centering
\includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_Significance_June}
\hspace{0.05\textwidth} \includegraphics[width=.45\textwidth]{FigsAndFiles/plotDiversity_TvT_Folded2_Significance_June}
\caption{Pairwise diversity surrounding synonymous and non-synonymous
  substitutions in maize and teosinte. \jri{a bit too much vertical white space} {\bf A:} Maize diversity; {\bf B:} Teosinte
diversity. Bootstrap-based 95\% confidence intervals are depicted via shading. Inset plots depict a larger range on the x-axis. \label{fig:hardSweeps}}
\end{figure*}

@@

\subsection{Pairwise diversity is heavily influenced by purifying selection}

Purifying selection refers to the situation where deleterious mutations arising in a population are continuously selected against.
When this form of selection is operating, it can serve to reduce genetic variability at linked neutral sites, a phenomenon often called background selection \cite{charlesworth1993}. Purifying and background selection lead to lower diversity within genes and other functional sites relative to neutral regions.
We investigated purifying selection in maize and teosinte by evaluating the average magnitude of reduced diversity within genes and recovery away from genes in both taxa.
To enable comparisons between maize and teosinte, we standardized by neutral levels within each taxa, where neutral levels were defined as mean diversity at positions at least 0.01 cM distal from genes. For both maize and teosinte, pairwise diversity was reduced within genes and a recovery of diversity away from genes was observed. Interestingly, a stronger reduction of diversity and slower recovery was observed for teosinte than for maize, implying that purifying selection has left a more pronounced signature on pairwise diversity in the teosinte genome (Figure \ref{fig:purify}).

@@

We used diffusion approximation as implemented in
$\delta\alpha\delta{i}$ \cite{gutenkunst2009} to find the domestication parameters that best explain
the joint site frequency spectrum of maize and teosinte.  The model we optimized began with an
equilibrium ancestral population of size $N_a$
splitting into separate maize and teosinte populations $T_B$ generations in the past. Moving forward in
time from the split, teosinte maintained the ancestral population size, while
maize experienced an immediate effective population size change to
$N_b$ individuals, followed by exponential growth to size
$N_m$. Additionally, since the split $M_{tm}$ individuals migrated
from teosinte into maize and $M_{mt}$ individuals migrated from maize
to teosinte.

@@

\begin{figure}[!htb]
%\vspace*{.05in}
\centering
\includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
\caption{Parameters of domestication as estimated by $\delta\alpha\delta{i}$. Notably,
  the maize effective population size ($N_e$) during the domestication
  bottleneck appears to have consisted of approximately 5.26\%
  the ancestral $N_e$, before recovering two at least 2.98
  times as large as the ancestral $N_e$. \label{fig:bottleneck} }
\end{figure}

@@

Ultimately, we have shown that purifying selection in maize has operated very differently than purifying selection in teosinte. This observation, along with the potential for soft sweeps, appear to explain the phenotypic divergence \jri{pretty hard to argue that purifying selection explains phenotypic divergence!} between maize and teosinte much more completely than does positive selection generating hard-sweeps at protein-coding mutations. Importantly, our estimation of the parameters of the maize domestication bottleneck contribute to the understanding of how the demography of crop domestication can impact crop diversity. The bottleneck-effects from a sudden collapse in population size have been well studied and are known to impact crops for thousands of generations. Complementing this knowledge, our results demonstrate that the rapid expansion experienced by many crops after domestication can also have a profound influence on patterns of diversity, and the effects of this expansion should be accounted for as important contributors to long-term evolution.

@@

\subsection{Plant materials}
Accessions studied were selected from the Maize HapMap2
panel \cite{chia2012, lemmon2014}. \jri{this needs slight clarification. cite chia for hapmap, but that we used higher-coverage of the teosinte from lemmon. if possible, include link to fastq or bamfiles used} Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results (Figure S5). Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion. 
Thirteen teosinte inbred lines, all members of  subspecies Z. \emph{mays}
ssp. \emph{parviglumis}, were utilized. Included lines are listed in (Table S1).
Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} (ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/) as described by \cite{chia2012}.\jri{did we use their bams? I thought we used bwa bams, which are not listed and weren't used for hapmap2. if that's the case, we need to mention this difference.}

\subsection{Interpolating genetic position}

Physical positions were converted to genetic positions by interpolating from the NAM genetic map \cite{glaubitz2014}, \jri{does jeff's paper really have the NAM map? I didn't think so and quick look didn't find it. i thought this came from wallace?} which provides a 1 cM resolution for physical to genetic conversion.
Within R \cite{R2014}, physical positions with corresponding genetic positions in the NAM map were used as anchors.
Physical positions in our dataset without corresponding genetic position were assigned genetic positions by scaling the anchored genetic positions according to the physical distance between the unlabeled position and the flanking anchors.

\subsection{Estimating the site frequency spectrum}
To estimate individual  and joint site frequency spectra (SFS) for maize and teosinte from inbred lines, each inbred individual was treated as representing a single haplotype from its population.
SFS were
separately computed for genic and intergenic regions, as well as for
the whole genome together. First, genic and intergenic regions were isolated using
the biomaRt package \cite{durinck2009,durinck2005} of R \cite{R2014}. Genic
regions were defined as DNA between the start and stop position of a \jri{you meant transcription start/stop not translation, right?}
gene, while intergenic regions were required to be at least 5kb up- or
down-stream from a gene start/stop. With regions defined, SFS were
estimated with ANGSD \cite{korneliussen2014}. Individual population SFS were
estimated using all positions observed in at least 80\% of the
individuals in the population, and joint SFS were estimated using all
positions observed in at least 80\% of individuals in both
populations. Individuals were
assumed to be fully inbred (-doSaf 2). Quality filters were
employed such that reads with quality score below 30 and bases with
quality score below 20 were discarded (-minMapq 30 and -minQ 20), as were reads that didn't map
uniquely (-uniqueOnly 0). Quality scores around indels were adjusted
as in Samtools (-baq 0). Genotype likelihoods were estimated using the samtools
method (-GL 1). Major and minor alleles were inferred from the data
(-doMaf 1). Because ANGSD cannot calculate a folded joint SFS, the
maize reference genome was used for polarization and then unfolded spectra
were folded using $\delta\alpha\delta{i}$ \cite{gutenkunst2009}.

\subsection{Demographic inference}
We used $\delta\alpha\delta{i}$ \cite{gutenkunst2009} to estimate the parameters of maize
domestication based on diffusion approximation of the 2-dimensional
SFS for maize and teosinte. Our first step was to specify a model, as
shown in (Figure \ref{fig:bottleneck}). In words, our model began with an ancestral teosinte
population with effective population size $N_a$ individuals. The
ancestral population split into
two distinct populations, one of maize and the other of teosinte,
$T_b$ generations in the past. The maize population immediately shrank
to size $N_b$, and then grew exponentially to size $N_{maize}$
today. From the split until today, the teosinte effective population
size remained constant. Each year, $M_{mt}$ individuals migrated from
maize to teosinte, and $M_{tm}$ individuals migrated from teosinte to
maize.The free parameters of our model were $N_a, N_b, N_{maize}, T_b,
M_{mt}, \text{and } M_{tm} $. We implimented 1,000 $\delta\alpha\delta{i}$
optimizations of this model using the ``$\delta\alpha\delta{i}$.Inference.optimize\_log\_fmin''
optimizer. In $\delta\alpha\delta{i}$ optimizations,
initial values for $N_b, N_{maize}, T_b,
M_{mt}, \text{and } M_{tm}$, respectively, were randomly perturbed up
to a factor of two from 0.02, 3, 0.04, 1e-5, and 1e-5. Lower
bounds were set at 1e-7, 1e-7,  0, 1e-10, and
1e-10, respecively, while upper bounds were respectively set at 2,
200, 1, .001, and .001. $N_a$ did not have an initial value, upper, or
lower bounds because it is estimated from equilibrium ancestral theta
($\theta = 4N_a\mu$), which $\delta\alpha\delta{i}$ determines automatically regardless
of model. Population sizes were converted from $\delta\alpha\delta{i}$'s
units using a mutation rate estimate of 3e-8
\cite{clark2005}. Among the 1,000 $\delta\alpha\delta{i}$ runs, parameters
from the run with the lowest log-likelihood were taken as estimates of
the parameters of the domestication process.

Recent population expansion, as is observed for maize, most influences
the abundance of rare alleles. However, our demographic parameter
estimation using $\delta\alpha\delta{i}$ suffered from a relatively small sample size of
23 maize individuals. Therefore, our approach likely underestimated
current maize effective population size. To address this, we utilized
a complimentary dataset of 4,021 maize landrace individuals collected
from across the Americas \textcolor{red}{(SeeDs reference here)} to
generate an independent estimate of
$N_{maize}$. These individuals were genotyped using GBS
\cite{elshire2011}, a protocol that suffers from a high rate of
missing data \cite{beissinger2013}. To minimize biases imposed from
missing data, we used R to isolate only those SNPs that
were observed in at least 1,500 individuals and subsequently projected the
SFS down to a sample of 500 individuals. Then, we utilized the singleton-based estimate of
theta ($\theta = 4N_e\mu = \frac{S}{L}$) to estimate
modern maize $N_e$ \cite{fu1993}, where $S$ is the total number of
singletons and $L$ is the total length of DNA sequenced. Based on the
GBS protocol implemented \cite{elshire2011}, $L$ is equal to the total
number of SNPs in the dataset.
